---
title: "Task 3"
author: "Lúa Arconada Manteca & Víctor Collado Fernández"
date: '27-09-2023'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

Create the following vectors

(a) $[10, 11, 12, \dots, 38]$

```{r exercise1 a, comment= ""}
a = c(10:38); a
```

(b) $[30, 29,28,\dots,1]$

```{r, comment= ""}
b = seq(30,1, -1); b
```

(c) $[1,2,3,4,3,2,1]$

```{r, comment= ""}
c = c(1,2,3,4,3,2,1); c
```

(d) $[2,4,6,\dots,16,18,20]$

```{r, comment= ""}
d = seq(2,20,2); d
```

(e) $[1,2,3,1,2,3,\dots,1,2,3]$. The pattern is repeated 10 times.

```{r, comment= ""}
e = rep(c(1:3), 10); e
```

(f) $[1,2,3,1,2,3,\dots,1]$. It is the vector of e) where the last two values have been removed.

```{r, comment= ""}
f = e[c(-29,-30)]; f
```

(g) ["label 1", "label 2", $\dots$, "label 30"]

```{r, comment= ""}
g = paste("label", c(1:30)); g
```

(h) ["label-1", "label-2", $\dots$, "label-30"]

```{r, comment= ""}
h = paste0("label", sep="-", c(1:30)); h
```

(i) $x^2e^x,\quad x=0.1,0.2,\dots,1$

```{r, comment= ""}
i = sapply(seq(0.1,1,0.1), function(x) x^(2)*exp(x)); i
```

# Exercise 2

Calculate the following sums with and without using loops.

$\sum_{j=5}^{23}(j^2+3*j^{0.5})$

```{r, comment= ""}
# With loop
s=0
for (j in 5:23) {
  s=s+j^2+3*j^(0.5)
}
s

# Without loop
j = 5:23
sum(j^(2)+outer(3,j^(0.5)))
```

$\sum_{i=1}^{18}\frac{1.3^i}{i}$

```{r, comment= ""}
# With loop
s=0
for (i in 1:18){
  s=s+(1.3^i)/i
}
s

# Without loop
i = 1:18
sum((1.3^(i))/i)
```

$\sum_{i=1}^{10}\sum_{j=1}^{6}\frac{i^4}{3+j}$

```{r, comment= ""}
# With loop
s=0
for (i in 1:10){
  for (j in 1:6){
    s=s+(i^4)/(3+j)
  }
}
s

# Without loop
i = 1:10 ; j = 1:6
sum(outer(i^4,3+j,"/"))
```

# Exercise 3

(a) What does the next code do?

```{r, comment= ""}
set.seed(75)
M = matrix(sample(1:10, size=60, replace = TRUE), nrow=6, ncol=10)
```

The code creates a matrix named M with 6 rows and 10 columns. It has 60 entries and they are chosen randomly in a range between 1 to 10 and the values can be repeated. The set.seed() command is used to obtain the same matrix independently of the random function that is used (sample).

(b) Find the number of entries in each row that are greater than 4.

```{r, comment= ""}
x=rep(0, 6)
for (row in 1:6) {
  s=0
  for (col in 1:10)
    if (M[row,col]>4) s=s+1
  x[row] = s 
}
paste ("row",1:6,"=",x[1:6])
```

(c) Replace the third column of the previous vector M by the sum of the second and third column.

```{r, comment= ""}
M[,3] = sum(M[,2]+M[,3]); M
```

# Exercise 4

Write a function which takes a single argument which is a matrix. The function must return a matrix which is the same as the function argument but every odd number is doubled.

```{r, comment= ""}
doub= function(A) {
  nrows = nrow(A)
  ncols = ncol(A)

  for (i in 1:nrows) {
    for (j in 1:ncols) {
      if (A[i, j] %% 2 != 0) {
        A[i, j] = A[i, j] * 2
      }
    }
  }
  return(A)
}


example <- matrix(1:10, nrow = 2)
example
example_modified=doub(example)
example_modified
```
